My title

### 2024-09-23 outputs/paper/paper.qmd
created 4 figures for the data section. llm usage updated

>>>>

```{r}

# install libraries 
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("ggplot2")
# install.packages("here")

# Load necessary libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)

# Check if the file exists and load it 
file_path <- here::here("inputs/data/analysis_data/neighborhood_analysis_data.csv")

if (file.exists(file_path)) {
  analysis_data <- read_csv(file_path)
  print("File found and loaded successfully!")
} else {
  stop("Error: File not found at the specified path.")
}

# Reshape the data to long format for easier plotting
long_data <- analysis_data %>%
  pivot_longer(cols = -`Neighbourhood Name`, 
               names_to = "Neighborhood", 
               values_to = "Value")

## Graphing visible minority data
# Define categories
visible_minority_categories <- c("South Asian", "Chinese", "Black", "Filipino",
                                 "Arab", "Latin American", "Southeast Asian",
                                 "West Asian", "Korean", "Japanese",
                                 "Visible minority, n.i.e.",
                                 "Multiple visible minorities",
                                 "Not a visible minority")

# Filter for visible minority categories
filtered_data <- long_data %>%
  filter(`Neighbourhood Name` %in% visible_minority_categories)

# Calculate total individuals for each neighborhood
total_individuals <- filtered_data %>%
  group_by(Neighborhood) %>%
  summarize(Total = sum(Value), .groups = "drop")  # Use .groups to avoid grouping issues

# Create the plot with number of individuals
visible_minorities_plot <- ggplot(filtered_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Visible Minority Categories Between Neighborhoods",
       x = "Visible Minority Categories",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "lightgreen", "Downtown Yonge East" = "forestgreen")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_data$Value), by = 2000))  # Adjust 'by' value as needed

# Display the plot
print(visible_minorities_plot)

## Graphing income data
# Define categories
income_categories <- c("Under $10,000", "$10,000 to $19,999",
                       "$20,000 to $29,999", "$30,000 to $39,999",
                       "$40,000 to $49,999", "$50,000 to $59,999",
                       "$60,000 to $69,999", "$70,000 to $79,999",
                       "$80,000 to $89,999", "$90,000 to $99,999",
                       "$100,000 to $149,999", "$150,000 and over")

# Convert the 'Neighbourhood Name' to a factor with specified levels for ordering
filtered_income_data$`Neighbourhood Name` <- factor(filtered_income_data$`Neighbourhood Name`, levels = income_categories)

# Create the plot with number of individuals for income
income_plot <- ggplot(filtered_income_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Income Categories Between Neighborhoods",
       x = "Income Categories",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "magenta", "Downtown Yonge East" = "darkmagenta")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_income_data$Value), by = 500))  # Adjust 'by' value for more tick marks

# Display the plot
print(income_plot)

## Graphing education data
# Define the ordered education categories
education_categories <- c("No certificate, diploma or degree",
                          "High school diploma or equivalency certificate", 
                          "Postsecondary certificate or diploma",
                          "Apprenticeship or trades certificate or diploma",
                          "Non-apprenticeship trades certificate or diploma",
                          "College, CEGEP or other non-university certificate or diploma",
                          "University certificate or diploma below bachelor level",
                          "Bachelor's degree",
                          "University certificate or diploma above bachelor level",
                          "Degree in medicine, dentistry, veterinary medicine or optometry",
                          "Master's degree", "Earned doctorate")

# Filter the long_data for education categories
filtered_education_data <- long_data %>%
  filter(`Neighbourhood Name` %in% education_categories)

# Convert 'Neighbourhood Name' to a factor for ordered plotting
filtered_education_data$`Neighbourhood Name` <- factor(filtered_education_data$`Neighbourhood Name`, levels = education_categories)

# Create the plot with number of individuals for education attainment
education_plot <- ggplot(filtered_education_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Education Attainment Between Neighborhoods",
       x = "Education Levels",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "skyblue", "Downtown Yonge East" = "blue")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_education_data$Value), by = 1000))  # Adjust 'by' value for more tick marks

# Display the plot
print(education_plot)

## Graphing knowledge of official language
# Define the categories for language knowledge
language_categories <- c("English only", "French only", 
                         "English and French", "Neither English nor French")

# Filter the long_data for language knowledge categories
filtered_language_data <- long_data %>%
  filter(`Neighbourhood Name` %in% language_categories)

# Convert 'Neighbourhood Name' to a factor for ordered plotting
filtered_language_data$`Neighbourhood Name` <- factor(filtered_language_data$`Neighbourhood Name`, levels = language_categories)

# Create the plot with number of individuals for knowledge of official languages
language_plot <- ggplot(filtered_language_data, aes(x = `Neighbourhood Name`, y = Value, fill = Neighborhood)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Knowledge of Official Languages Between Neighborhoods",
       x = "Knowledge of Official Languages",
       y = "Number of Individuals") +
  theme_minimal() +
  scale_fill_manual(values = c("Rosedale-Moore Park" = "yellow", "Downtown Yonge East" = "gold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, max(filtered_language_data$Value), by = 2000))  # Adjust 'by' value for more tick marks

# Display the plot
print(language_plot)

```
